---
title: "Stealing Discovery Analysis"
author: "Dusty Turner"
date: "June 13, 2019"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
```

## Data

1. The data we are analyzing below come from the following website: [Baseball Savant]("https://baseballsavant.mlb.com/statcast_search")

2. I learned by trial and error that you can only download 40,000 rows of data at a time.  You'll have to add it together to do more stuff.  

## Read in Data and Set Options

```{r}
library(tidyverse)
library(stringr)
options(tibble.print_min = 100)
data = read_csv("savant_data_stealing.csv",guess_max = 10000) %>% janitor::clean_names()
# data = read_csv("savant_data_stealing_everypitch.csv",guess_max = 10000) %>% janitor::clean_names()
```

The data has the following columns available:

```{r}
names(data)
```

The dates contain the following range:

```{r}
data$game_date %>% range()
```

Based on the way I requested the data, the data only contains plate appearances where there was at least one runner on base

## Determining Who Stole Second

My apologies for not commenting this well.  I recomend running this line by line to determing what is happening.

```{r}
##  
stolesecond =
data %>% 
  mutate(pitchID = row_number()) %>%
  # select(pitchID, events,des,pitch_type,release_speed,batter,at_bat_number,on_1b,on_2b,on_3b) %>%
  select(pitchID, batter, pitcher,balls, strikes,at_bat_number,on_1b,on_2b,on_3b) %>%
  mutate_at(c("on_1b","on_2b","on_3b"), as.numeric) %>%
  mutate(on_1b = na_if(on_1b,"null"), on_2b = na_if(on_2b,"null"), on_3b = na_if(on_3b,"null")) %>%
  group_by(batter,pitcher,at_bat_number) %>%
  arrange(batter) %>%
  filter(sum(is.na(on_2b))>1&&sum(!is.na(on_2b))>1) %>%
  group_by(batter,pitcher,at_bat_number,on_2b) %>%
  mutate(groupID = row_number())  %>%
  ungroup() %>%
  filter(!is.na(on_1b)|!is.na(on_2b)) %>%
  filter(!is.na(on_2b)) %>%
  group_by(batter,pitcher,at_bat_number,on_2b) %>%
  filter(groupID == max(groupID)) %>% ## this is the pitch he stole from 1st to second 
  ungroup() %>%
  select(pitchID,batter) %>%
  rename(stole2nd = batter) 
```

Look at it:

```{r}
stolesecond %>% head
stolesecond %>% count(stole2nd,sort = TRUE) %>% top_n(10)
```

Now that we've discover who has stole second, lets add that data back into the main dataset.

```{r}
secondseal =
data %>%
  # select(game_date,batter, des,pitcher,balls, strikes,outs_when_up,at_bat_number,on_1b,on_2b,on_3b, release_speed,pitch_type) %>%
  mutate(pitchID = row_number(), des = na_if(des,"null")) %>%
  left_join(stolesecond) %>%
  mutate(stole2nd = ifelse(is.na(stole2nd), "No","Yes")) %>% 
  mutate(stole2ndcaught = ifelse(str_detect(des,"caught")&str_detect(des,"2nd base"), "Yes", "No")) %>% 
  mutate(stealattempt = ifelse(stole2nd=="Yes"|stole2ndcaught=="Yes","Yes","No")) %>%
  mutate(release_speed = as.numeric(release_speed)) %>% filter(stealattempt == "Yes")
  # select(pitchID,des, stole2nd,stole2ndcaught,stealattempt) 
```

Look at data:

```{r eval=FALSE, message=FALSE, warning=FALSE, include=TRUE, paged.print=FALSE}
secondseal %>% head
```

\newpage

## Visualizations


```{r}
secondseal %>% 
  ggplot(aes(x=release_speed, fill = stole2nd)) + geom_density(, alpha = .5) +
  labs(title = "Successful Steals at Different Release Speeds")
```

```{r}
secondseal %>% 
  # select(pitch_type,release_speed,stole2nd) %>%
  select(release_speed,stole2nd, outs_when_up) %>%
  gather(key, other,-release_speed,-stole2nd) %>%
  ggplot(aes(x=release_speed, fill = stole2nd)) + geom_density(, alpha = .5) +
  facet_wrap(~other) +
  labs(title = "Successful Steals at Different Selease Speeds", subtitle = "By Number of Outs")
```

```{r}
secondseal %>% 
  select(pitch_type,release_speed,stole2nd) %>%
  # select(release_speed,stole2nd, outs_when_up) %>%
  gather(key, other,-release_speed,-stole2nd) %>%
  ggplot(aes(x=release_speed, fill = stole2nd)) + geom_density(, alpha = .5) +
  facet_wrap(~other) +
  labs(title = "Successful Steals at Different Selease Speeds", subtitle = "By Pitch Type")
```

## Stealing Third Attempts

Not currently correct

```{r echo=FALSE, eval=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
# stolethird =
# data %>% 
#   mutate(pitchID = row_number()) %>%
#   # select(pitchID, events,des,pitch_type,release_speed,batter,at_bat_number,on_1b,on_2b,on_3b) %>%
#   select(pitchID, batter, pitcher,balls, strikes,at_bat_number,on_2b,on_3b) %>%
#   mutate_at(c("on_2b","on_3b"), as.numeric) %>%
#   mutate(on_2b = na_if(on_2b,"null"), on_3b = na_if(on_3b,"null")) %>%
#   group_by(batter,pitcher,at_bat_number) %>%
#   arrange(batter) %>%
#   # filter(sum(is.na(on_1b))>0&&sum(!is.na(on_1b))>0)
#   filter(sum(is.na(on_2b))>0) %>%
#   filter(sum(!is.na(on_2b))>0) %>%
#   # group_by(on_2b,on_3b) %>%
#   ungroup() %>%
#   mutate(groupnumber = row_number()) %>%
#   group_by(batter,pitcher,at_bat_number) %>%
#   mutate(help = sum(!is.na(on_3b))) %>%
#   filter(help!=0) %>%
#   filter(sum(is.na(on_3b))!=0) %>%
#   group_by(batter,pitcher,at_bat_number,on_3b) %>%
#   filter(is.na(on_2b)|is.na(on_3b)) %>%
#   group_by(batter,pitcher,at_bat_number) %>%
#   mutate(row_number()) %>%
#   ungroup() %>%
#   filter(!is.na(on_2b))

```

\newpage


## A Look at Other Batting Stuff

```{r}
texas = read_csv("texas_batters.csv") %>% janitor::clean_names() %>%
  mutate(release_speed = as.double(release_speed), 
         launch_speed = as.double(launch_speed), 
         launch_angle = as.double(launch_angle),
         # launch_angle = as.double(launch_angle),
         hit_distance_sc = as.double(hit_distance_sc),
         hc_x = as.double(hc_x),
         hc_y = as.double(hc_y))


texas %>% 
  select(player_name,release_speed,launch_speed,pitch_type,launch_angle,hit_distance_sc) %>%
  filter(player_name=="Elvis Andrus") %>%
  mutate(pitch_type = fct_lump(pitch_type)) %>%
  ggplot(aes(x=release_speed, y=hit_distance_sc)) +
  geom_point(aes(color = launch_angle)) +
  facet_wrap(~pitch_type) +
  labs(title = "Hit Distance vs Pitch Speed", x="Pitch Speed", y = "Hit Distance", 
       color = "Launch Angle", subtitle = "Elvis Andrus")


```



```{r}
texas %>% 
  select(player_name,release_speed,launch_speed,pitch_type,
         launch_angle,hit_distance_sc, hc_x,hc_y,stand,of_fielding_alignment) %>%
  # filter(player_name=="Elvis Andrus") %>%
  mutate(of_fielding_alignment = fct_lump(of_fielding_alignment)) %>%
  ggplot(aes(x=hc_x, y=hc_y)) +
  geom_point(aes(color = launch_angle)) +
  facet_grid(of_fielding_alignment~stand) +
  labs(title = "Hit Distance vs Pitch Speed", x="Pitch Speed", y = "Hit Distance", 
       color = "Launch Angle", subtitle = "Elvis Andrus")

```